---
title: "Regression Model Project"
author: "Thibault Dubois"
date: "18 mars 2018"
output: md_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

#Executive Summary

You work for Motor Trend, a magazine about the automobile industry. Looking at a data set of a collection of cars, they are interested in exploring the relationship between a set of variables and miles per gallon (MPG) (outcome). They are particularly interested in the following two questions:

*Question 1:*
“Is an automatic or manual transmission better for MPG”

*Question 2:* 
"Quantify the MPG difference between automatic and manual transmissions"

After analyzing the data, we arrive to the following conclusions:

*Answer 1:*
Manual transmissions perform better than automatic transmissions by 7.25MPG, however this single factor only accounts for 36% of the explanation

*Answer 2:*
When measuring MPG, manual transmissions provide an additional 1.48MPG of performance over automatic transmissions when taking into account three additonal explanatory variables (cylinders, horsepower & weight), these additional factors account for 85% of the explanation

#Exploratory Data Analysis

###Loading the data and observing the data set:
```{r}
library(datasets)
data(mtcars)
```

```{r}
head(mtcars, 3)
```

```{r}
str(mtcars)
```

###Further exploration of the data set

```{r}
summary(mtcars$mpg)
```
We see that the median and the mean of mpg are close to each other, suggesting a normal distribution. Plotting the data will give a better view on the data. 

```{r}
library(ggplot2)
ggplot(data = mtcars, aes(x = mpg))+geom_histogram(binwidth = 3)
```
This graph suggests that the data for mpg is slightly rightly skewed. 

We would also like to see how mpg changes when controlled for autmatic transmission versus manual transmission. Boxplots are a good way to show the relationships between numerical and categorical variables
```{r}
#Visualization ~ Automatic vs Manual Transmission:
library(ggplot2)
mtcars$am <- as.factor(mtcars$am)
Trans_Type <- ggplot(aes(x=am, y=mpg), data=mtcars) + geom_boxplot(aes(fill=am))
Trans_Type <- Trans_Type + labs(title = "Automatic vs Manual Transmission")
Trans_Type <- Trans_Type + xlab("Transmission Type")
Trans_Type <- Trans_Type + ylab("MPG")
Trans_Type <- Trans_Type + labs(fill = "Legend (0=AT, 1=MT)")
Trans_Type
```
This boxplot is a first indication of the drivers behind mpg. We clearly see that automatic transmissions are consume on average less than manual transmissions. 

Lets review the summary statistics but this time split between the two groups.

```{r}
Trans_Stats = split(mtcars$mpg, mtcars$am)
```

The mean:
```{r}
sapply(Trans_Stats, mean)
```

The standard deviation: 
```{r}
sapply(Trans_Stats, sd)
```

The maximum and the minimum
```{r}
sapply(Trans_Stats, range)
```
Looking at the summary statistics


#Hypothesis testing
We would like to formally test wether there is a significant difference between automatic and manual transmissions in terms of mpg. In order to do this the following t-test has been executed:

```{r}
Auto_trans <- mtcars[mtcars$am == "0",]
Man_trans <- mtcars[mtcars$am == "1",]
t.test(Auto_trans$mpg, Man_trans$mpg)
```
P-value being smaller than 5% we can safely conclude that there is a signficant difference between manual transmissions and automatic transmissions. Automatic transmissions consume significantly less than manual transmissions. Now we would like to exactly measure this difference through linear regressions.  

#Linear Regression Model

###Single regressor model
First, I conduct a regression with only the am variable:
```{r}
Model1 <- lm(mpg ~ am, data = mtcars)
summary(Model1)
```
*Interpreation*
Automatic transmission: the mean mpg consumed is equal to 17.147 mpg
Manual transmission: the mean mpg consumed is qual to 17.147 mpg + 7.245 mpg = 24.392 mpg.
In other words manual tranmission consumed on average 7.245 mpg *more* than automatic transmission. 

However, when we look at adjusted R² we see that the model only explains 33.85% of the variation in the observations. We can try to get a more accurate and parsimonuous model by adding extra variables. 


We find the same results (with a slightly different interpreation) when exluding the intercept. Here we see the means when compared to 0.
```{r}
Model1.1 <- lm(mpg ~ am - 1, data = mtcars)
summary(Model1.1)
```

###Multivariate regressor model
This model looks at 4 additional regressors:
1. Automatic versus manual transmission
2. Cylinder type
3. Horsepower
4. Weight
```{r}
Model2 <- lm(mpg~am + cyl + hp + wt, data = mtcars)
summary(Model2)
```
*Interpretation*
F-statistic:
The model as a whole is significant with a very high F-statistic and a low p-value.

Individual P-values:
We see that am1 is still highly significant with the intercept however we the impact of having a automatic transmission on mpg is considerably lower (from 7.245 in model 1 to 1.478 model 2). 

Interesting result is weight. It appears that more weight decreases mpg which might sound counter intuitive. Weight is also signficant 

Model accuracy:
Model 2 is more accurate in explaing the variance in observation with an adjusted R² of 82.67% (as opposed to 33.85%). 

We can make an additional anova test to se wether the models are significantly different from each other. In other words, we test wether adding the extra variables produces a different model. 
```{r}
anova(Model1, Model2)
```
*Interpretation*
We reject the null hypothesis of the two models being equal. 

#Appendix 

This plot helps to detect wether certain variables are correlated with each other. Correlation between the regressors can bias the predictors.  
```{r}
pairs(mpg ~ ., data = mtcars)
```
We see for instance that weight and horsepower show a positive correlation. 

This plot summarises the most important tests for a regression model:
- Linearity
- Normal distribution
```{r}
par(mfrow = c(2,2))
plot(Model2)
```
*Interpretation*
- We see no real pattern in the residuals so the data is homoskedastic
- The plots follow more or less the line, suggesting the model is linear but skewed to the right
- Some outliers have influence and leverage such as the Toyota Corolla
